Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-22 Draft.Rmd
Update Draft.Rmd
lin regr done

  Air pollution has consequences for everyone, most especially for those with pre-existing conditions. Air pollution is measured in tiny particles or droplets in the air that are two and one-half microns or less in width called PM2.5. So being able to predict the average PM2.5 concentrations ($\mu$g/m$^3$) will allow us to keep people most vulnerable safe as we look for and implement changes to improve the air quality in the region.
We chose the predictors by …
For our four models, we chose linear regression, k-nearest neighbors, xgboosting, and random forest. Linear regression models a linear relationship between the PM2.5 value in the atmosphere and the predictors. K-nearest neighbors (kNN) returns predicted PM2.5 values based on their neighboring data points. INSERT BOOSTING MODEL SENTENCE. Random forest makes an averaged prediction from a collection of independent decision trees. We predict that the RMSE performance of the bootstrapping model will be the best, with a value of around 0.5.

```{r}
library(tidyverse)
library(tidymodels)
library(broom)
library(caret)
library(plotROC)
library(randomForest)

@@

# Visualize linear relationship between predictors (lat, lon, state) and value
origin %>% ggplot(aes(x=lon, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration (mu g/m^3)", x="Longitude (degrees)", y="PM2.5")

origin %>% ggplot(aes(x=lat, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration (mu g/m^3)", x="Latitude (degrees)", y="PM2.5")

origin %>% ggplot(aes(x=value, fill=state)) + geom_histogram() + facet_wrap(~state) + 
  labs(title="PM2.5 Concentration (mu g/m^3)", x="State", y="PM2.5") 
```

## Predictor/Feature Selection
```{r}
# Use Random Forest for variable selection
rfmodel = randomForest(value~., data=train[,2:50])
import=as.data.frame(importance(rfmodel)) |> arrange(desc(IncNodePurity))
import=tibble::rownames_to_column(import, 'variables')
# Now from this list, you can select the variables. Mostly, the top 20 variables are what we select. However, this is just an indicative value and may vary based upon the problem.

@@

# fit randfom forest and get var. importance based on mean decr. in accuracy
set.seed(123)
library(party)
train_num = train[,2:50] |> dplyr::select(where(is.numeric))
cf1 = cforest(value ~ ., data=train_num, control=cforest_unbiased(mtry=2, ntree=50))
# adj for corr. btw predictors
varimp(cf1, conditional=TRUE)
# more robust towards class imbalance
varimpAUC(cf1)

# Scale dataset
# Covar. matrix
# Eigenval.s and eigenvec.s
# PCA
# Data transformation in new dimensional space
```

## Models 
```{r Lin Regr.}
model1 <- linear_reg() %>% 
    set_engine("lm") %>% 
    set_mode("regression")

imp = varImp(model1, scale=FALSE)

fit1 = model1 |> fit(value ~ ., data=train_scaled)
x=summary(fit1$fit)$coefficients

tidy(fit1)
glance(fit1)

wf1 <- workflow() %>% 
    add_recipe(rec1) %>% 
    add_model(model1)

res1 = wf1 |> last_fit(split=dat_split)
res1 |> collect_metrics()

library(MASS)
fit <- lm(value ~ .,data=train_scaled)
step <- stepAIC(fit, direction="forward")
step$anova

```

```{r Linear Regression}
set.seed(123)
# Create the recipe for all models
rec <- train %>% 
    recipe(value ~ )

@@

wf1 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model1)
# wf1

res1 <- fit(wf1, data = train)

@@

# Check linear regression model performance using cross-validation

folds1 <- vfold_cv(train, v = 10)
folds1
res1 <- fit_resamples(wf1, resamples = folds1)
res1 %>% 
    collect_metrics()
```

@@

# Scale the training data for k-NN model in wrangling

>>>>

  Air pollution has consequences for everyone, most especially for those with pre-existing conditions. Air pollution is measured in tiny particles or droplets in the air that are two and one-half microns or less in width called PM2.5. So being able to predict the average PM2.5 concentrations ($\mu$g/m$^3$) will allow us to keep people most vulnerable safe as we look for and implement changes to improve the air quality in the region.
We chose the predictors by …
For our four models, we chose linear regression, k-nearest neighbors, xgboosting, and random forest. Linear regression models a linear relationship between the PM2.5 value in the atmosphere and the predictors. K-nearest neighbors (kNN) returns predicted PM2.5 values based on their neighboring data points. **INSERT BOOSTING MODEL SENTENCE.** Random forest makes an averaged prediction from a collection of independent decision trees. We predict that the RMSE performance of the bootstrapping model will be the best, with a value of around 0.5.

```{r}
library(tidyverse)
library(tidymodels)
library(broom)
library(xgboost)
library(caret)
library(plotROC)
library(randomForest)
library(factoextra)
library(MASS)

@@

# Visualize linear relationship between predictors (lat, lon, state) and value
origin %>% ggplot(aes(x=lon, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration vs Longitude", x="Longitude (Degrees)", y="PM2.5 (mu g/m^3)")

origin %>% ggplot(aes(x=lat, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration vs Latitude", x="Latitude (Degrees)", y="PM2.5 (mu g/m^3)")

origin %>% ggplot(aes(x=value, fill=state)) + geom_histogram() + facet_wrap(~state) + 
  labs(title="PM2.5 Concentration", x="PM2.5 (mu g/m^3)", y="Frequency (Monitors)") + theme(legend.position='none')
```

## Predictor/Feature Selection
```{r}
# Use Random Forest for variable selection
rfmodel = randomForest(value~., data=train[,2:50])
import=as.data.frame(importance(rfmodel)) |> arrange(desc(IncNodePurity))
import=tibble::rownames_to_column(import, 'variables')

@@

# fit randfom forest and get var. importance based on mean decr. in accuracy
set.seed(123)
library(party)
train_num = train[,2:50] |> dplyr::select(where(is.numeric))
cf1 = cforest(value ~ ., data=train_num, control=cforest_unbiased(mtry=2, ntree=50))
# adj for corr. btw predictors
varimp(cf1, conditional=TRUE)
# more robust towards class imbalance
varimpAUC(cf1)
```

## Models 
```{r Linear}
set.seed(123)
# Fit model and get summary
fit_lm = lm(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data=train)
step <- stepAIC(fit_lm, direction="backward", trace=TRUE)
# Drops county_area, popdens_county, and log_nei_2008_pm25_sum_10000

# Create the recipe for all models
rec1 <- train %>% 
    recipe(value ~ CMAQ+lat+lon+aod+zcta+log_nei_2008_pm25_sum_10000)

@@

wf1 <- workflow() %>% 
    add_recipe(rec1) %>% 
    add_model(model1)
# wf1

res1 <- wf1 |> parsnip::fit(data = train)

@@

# Check linear regression model performance using cross-validation
folds1 <- vfold_cv(train, v = 10)
# folds1
res1 <- fit_resamples(wf1, resamples = folds1)
res1 %>% 
    collect_metrics()
```

@@

# Scale the training data for k-NN model in wrangling
train_scaled = train[,3:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) 
#%>% cbind(state=train$state,county=train$county,city=train$city)
